Certified Knowledge Compilation \\ with Application to Verified Model Counting

### 2023-03-18 doc/lean.tex
Polished main part of paper

Our specifications are built on a generic library for the syntax and semantics of
propositional logic. We define the data type \lstinline{PropForm Var} of propositional
formulas with variables \lstinline{Var}, which are positive natural numbers,
corresponding to the indexing of variables in the CNF file.
It is convenient to take a truth assignment \lstinline{PropAssignment Var} to a function that
assigns a Boolean value to every assignment, and later restrict attention to subsets of the
variables.

@@

The goal of the proof checker is to construct a POG that is equivalent to the input CNF.
Each element of a POG  is either a variable, a binary disjunction,
or an arbitrary conjunction:
\begin{lstlisting}
inductive PogElt where
  | var  : Var → PogElt
  | disj : Var → ILit → ILit → PogElt
  | conj : Var → Array ILit → PogElt
\end{lstlisting}
In the first case, the argument \lstinline{x} in the expression
\lstinline{var x} is the index
of the variable; in \lstinline{disj x left right} and \lstinline{conj x args}
it is the definition number in the CRAT file. Note that representing edges as literals allows us to negate the arguments to \lstinline{disj} and \lstinline{conj}.
A \lstinline{Pog} is then simply an array of \lstinline{PogElt}.
For each POG \lstinline{P} and literal \lstinline{l}, we define
\lstinline{P.toPropForm l} to be the
propositional formula that arises from interpreting \lstinline{l} as a propositional
formula, unfolding all the defined conjunctions and disjunctions.
A partitioned-operation formula, as defined in \cref{section:logical:foundations}, is expressed as:
\begin{lstlisting}
def partitioned : PropForm Var → Prop
 | tr | fls | var _ => True
 | neg φ    => φ.partitioned
 | disj φ ψ => φ.partitioned ∧ ψ.partitioned ∧ ∀ τ, ¬(φ.eval τ ∧ ψ.eval τ)
 | conj φ ψ => φ.partitioned ∧ ψ.partitioned ∧ (φ.vars ∩ ψ.vars = ∅)
\end{lstlisting}

>>>>

Our specifications are built on a generic library for the syntax and semantics of
propositional logic. We define the data type \lstinline{PropForm Var} of propositional
formulas with variables \lstinline{Var}, which are positive natural numbers,
corresponding to the indexing of variables in the CNF and CRAT files.
It is convenient to take a truth assignment \lstinline{PropAssignment Var} to a function that
assigns a Boolean value to every assignment, and later restrict attention to subsets of the
variables.

@@

The goal of the proof checker is to construct a POG that is equivalent to the input CNF.
Each element of a POG  is either a variable, a binary disjunction (sum),
or an arbitrary conjunction (product):
\begin{lstlisting}
inductive PogElt where
  | var  : Var → PogElt
  | disj : Var → ILit → ILit → PogElt
  | conj : Var → Array ILit → PogElt
\end{lstlisting}
In the first case, the argument \lstinline{x} in the expression
\lstinline{var x} is the index
of the variable; in \lstinline{disj x left right} and \lstinline{conj x args}
it is the definition number in the CRAT file. Note that representing edges as literals allows us to negate the arguments to \lstinline{disj} and \lstinline{conj}.
A \lstinline{Pog} is then simply an array of \lstinline{PogElt}.
For each POG \lstinline{P} and literal \lstinline{l}, we define
\lstinline{P.toPropForm l} to be the
propositional formula that arises from interpreting \lstinline{l} as a propositional
formula, unfolding all the defined conjunctions and disjunctions.
A partitioned-operation formula, as defined in \cref{section:logical:foundations}, is expressed as:
\begin{lstlisting}
def partitioned : PropForm Var → Prop
 | tr | fls | var _ => True
 | neg φ    => φ.partitioned
 | disj φ ψ => φ.partitioned ∧ ψ.partitioned ∧ ∀ τ, ¬(φ.eval τ ∧ ψ.eval τ)
 | conj φ ψ => φ.partitioned ∧ ψ.partitioned ∧ (φ.vars ∩ ψ.vars = ∅)
\end{lstlisting}
